Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-12 paper/structureScene.tex
Edit results section

In the experiments reported in this paper, the DCASE2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE2016 datasets  \cite{Mesaros2016_EUSIPCO} are considered.

The DCASE2013  dataset consist of 2 parts, namely public and private, each made up of 100 30sec recordings of various acoustic scenes. The 100 recordings are equally divided into 10 acoustic scene classes. To build the DCASE2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio. No
systematic variations in the recordings covaried with scene
type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a consequence, DCASE2013 dataset has interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.

In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieves  between $84\%$ and $87\%$ on DCASE2013 depending on the prior knowledge used during training \cite{phan2016label}.

@@

The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a fivefold cross validation scheme. The folds used in this paper are the same as the ones used during the challenge.

@@

Experiments are carried out using scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features. For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$. Indeed, $3$ seconds are discarded at the beginning and at the end of the scene so as to avoid boundary artifacts.  Experiments are conducted with and without the logarithmic compression (see Section \ref{sec:logcomp}).

MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range. Different ranks of coefficients were tested to compute the MFCCs, results here are only reported for the best setting, which consisted of $40$ MFCCs, including the first coefficient related to the average energy. This setting gives us $600$ feature vectors per scene. To get a more robust representation of lower dimensionality a pooling step is performed on the MFCCs: each recording is divided into $n$ non-overlapping 250 seconds long ``texture'' windows \cite{1021072}, achieved a comparable number of vectors than while using the scattering ones.

@@

For ASSR, evaluation is performed on the private part of the DCASE2013 dataset. The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items. The $p@k$ is computed for $k=\{1,\ldots,9\}$, since each class only has $10$ items. Note that a $p@1$ is equivalent to the classification accuracy obtained by choosing the label of the closest neighbor for a given item. For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameter $N$ of the RBF kernels (see Eq. \ref{eq:kc}) leading to the best $p@9$.

@@

For ASC, the systems are again evaluated on the private part of DCASE2013, but here with five-fold stratified cross validation. The folds are identical to those used in the original challenge. Optimal parameters (see above) are this time learned on the public part of the DCASE2013 dataset, also using five-fold stratified cross validation.

@@

For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}. For both the \emph{early} and \emph{late} approaches, a linear kernel as well as a RBF kernel are used. As in the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}. All feature vectors, be they frames, texture windows or centroids, are standardized prior to be given to the classifier. The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).

Clustering is done using KMeans$++$ \cite{arthur2007k}, an augmented version of K-means with a robust initialization procedure. Three numbers of clusters are tested: 8, 16 and 32.

@@

Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach. Overall, log-compressed scattering coefficients systematically outperform MFCCs. This is to be expected since the scattering coefficients capture larger-scale modulations as opposed to the MFCCs which only describe the short-time spectral envelope.

@@

For the scattering transform, \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes. However, it is worth noticing that \emph{ob-a} perform equivalently to \emph{early}, which shows that the discriminant information is destroyed by averaging the contributions from all centroids. To take advantage of an object-based representation, we need to select certain representative centroids when comparing objects. Furthermore, it appears that \emph{ob-c} is be able to better characterize the classes than \emph{ob-w}. This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution. Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames. On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.

The same observations are made for the MFCCs for a $k\leq5$. However for a rank $k$ greater than 5, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}. The latter even outperform \emph{ob-a} for $k\geq5$. From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class. This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.

@@

\subsubsection*{Early \vs late integration}
Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to segregate distinct acoustic events beyond the scale of about one second.
Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none of them managed to outperform the baseline strategy in temporal integration, namely majority voting. Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, and reaches a accuracy of $66\% \pm0.7$. This might be due to the fact that the discriminability gained by the object-based approach is already gained by the classifiers during the training and the fact that considering the object-base approach resort to considering an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size, see discussion in Section \ref{sec:eili}.

@@

Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
This transformation improves the discriminative power of support vector machines (SVM) trained on scattering coefficients, whatever be the chosen temporal integration strategy (early or late) or the inner product kernel (linear or Gaussian RBF).
Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.

@@

\subsubsection*{Participation to the DCASE 2016 challenge}
In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge, whose results are forthcoming.
This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

Our results on the development dataset are reported in Table \ref{table:dcase2016}, in comparison with a baseline of MFCCs coefficients and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model. The baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
The average accuracy across classes is brought from $70.8\%$ to $79.4\%$.
Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
The results on the evaluation dataset are awaiting publication by the DCASE organizing committee, and will be disclosed in August 2016.

>>>>

In the experiments reported in this paper, we consider the DCASE 2013 \cite{giannoulis2013database, 7100934} and the public part of the DCASE 2016 datasets  \cite{Mesaros2016_EUSIPCO}.

The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes. The $100$ recordings are evenly divided into $10$ acoustic scene classes. To build the DCASE 2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio. No
systematic variations in the recordings covaried with scene
type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a consequence, DCASE2013 dataset has interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.

In addition, it is still a challenging dataset, with a state-of-the-art system based on feature design and SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieves between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.

@@

The dataset is split into a public dataset that can used for optimizing the ASC system and a private dataset used for computing the resulting accuracy using a five-fold cross validation scheme. The folds used in this paper are the same as the ones used during the challenge.

@@

Experiments are carried out using scattering coefficients as well as mel-frequency cepstral coefficients (MFCCs) as features. For the scattering transform, each $30$-second scene is described by $128$ vectors computed with half-overlapping windows $\boldsymbol{\phi}(t)$ of duration $T=372\,\mathrm{ms}$, for a total of $24\,\mathrm{s}$. Indeed, $3$ seconds are discarded at the beginning and end of the scene to avoid boundary artifacts. Experiments are conducted with and without logarithmic compression (see Section \ref{sec:logcomp}).

MFCCs are computed for windows of $50\,\mathrm{ms}$ and hops of $25\,\mathrm{ms}$, with full frequency range. Different ranks of coefficients were tested to compute the MFCCs, results here are only reported for the best setting, which consisted of $40$ MFCCs, including the first coefficient related to the average energy. This setting gives us $600$ feature vectors per scene. To get a more robust representation, a pooling step is performed on the MFCCs: each recording is divided into $250$-second long non-overlapping ``texture'' windows \cite{1021072}, over which the feature vectors are averaged. This yields number of vectors comparable to that of the scattering transform.

@@

For ASSR, evaluation is performed on the private part of the DCASE 2013 dataset. The metric used is the precision at rank $k$ ($p@k$), which is computed by taking a query item and counting the number of items of the same class within the $k$ closest neighbors, and then averaging over all query items. The $p@k$ is computed for $k=\lbrace 1,\ldots,9\rbrace$, since each class only has $10$ items. Note that a $p@1$ is equivalent to the classification accuracy obtained by the classifier which chooeses the label of the closest neighbor for a given item. For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $M$ for the \emph{ob} approaches and the scaling parameter $k$ of the RBF kernels (see Eq. \ref{eq:kc}) leading to the best $p@9$.

@@

For ASC, the systems are again evaluated on the private part of DCASE 2013, but here with five-fold stratified cross validation. The folds are identical to those used in the original challenge. Optimal parameters (see above) are this time learned on the public part of the DCASE 2013 dataset, also using five-fold stratified cross validation.

@@

For the \emph{ob} approaches, we use the precomputed kernels described in Section \ref{sec:object}. For both the \emph{early} and \emph{late} approaches, a linear kernel as well as an RBF kernel are used. As with the RBF kernels used in the \emph{ob} approaches, the RBF kernel in the SVM is scaled locally using the method proposed in \cite{selfTuneManor2004}. All feature vectors, be they frames, texture windows or centroids, are standardized prior to be given to the classifier. The standardization is computed with respect to the train/test splits of the cross validation scheme (see Section \ref{sec:stand}).

Clustering is done using $k$-means$++$ \cite{arthur2007k}, an augmented version of $k$-means with a more robust initialization procedure. Three numbers of clusters are tested: $8$, $16$ and $32$.

@@

Irrespective of the rank $k$ considered, best result is achieved for the scattering transform with logarithmic compression using the \emph{ob-c} approach. Overall, log-compressed scattering coefficients systematically outperform MFCCs. This is to be expected since the scattering coefficients capture larger-scale modulations, as opposed to MFCCs which only describe the short-time spectral envelope.

@@

For the scattering transform, both \emph{ob-c} and \emph{ob-w} outperform \emph{early}, thus confirming the benefits of using an object-based representation to refine the similarity measures between the scenes. However, it is worth noticing that \emph{ob-a} performs comparably to \emph{early}, showing that the discriminant information is destroyed by averaging the contributions from all centroids. To take advantage of an object-based representation, we need to select certain representative centroids when comparing objects. Furthermore, it appears that \emph{ob-c} is better able to characterize the classes than \emph{ob-w}. This last observation suggests that weighting a centroid according to the number of frames it contains may prove to be a limited solution. Indeed, nothing a priori indicates that the discriminant information between two scenes lays within the majority of their frames. On the contrary, two similar environments may shared a lot of similar sound sources with only a few sources discriminating between them.

The same observations are made for the MFCCs for a $k\leq5$. However for a rank $k$ greater than $5$, there is no clear evidence that \emph{ob-c} and \emph{ob-w} approaches improve the results compared to \emph{early}. The latter even outperforms \emph{ob-a} for $k\geq5$.

From this it appears that, although the \emph{ob} approaches are able to better retrieve the similarities between close scenes compared to \emph{early}, they are of no help when considering distant scenes of the same class. This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making a centroid-based description useless, or even counterproductive when the centroids are not weighted/selected.

@@

\subsubsection*{Early \vs late integration}
Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to separately characterize distinct acoustic events beyond the scale of about one second.
Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.
Surprisingly, although we ran an extensive number of experiments to formulate object-based clustering into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting. Among object-based methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$.

%\ml{explain why ?}

@@

Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.

@@

\subsubsection*{Participation to the DCASE 2016 challenge}
In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model. This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016}.
Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
